Modeling the effect of aneuploidy on cancer evolution

### 2023-01-31 EvolutionaryRescue1.tex
Add files via upload

\draw [thick](0.15+10,-1.25+2-4) -- (0.15+10,-2.75+2-4);
\draw [thick](-0.25+10,-1.25+2-4) -- (-0.25+10,-2.75+2-4);
\draw [,color=blue,very thick](-0.7+10,-1.25+2-4) -- (-0.7+10,-2.75+2-4);

\draw [thick](0.05+5,-1.25+2-8) -- (0.05+5,-2.75+2-8);
\draw [,color=blue,very thick](-0.55+5,-1.25+2-8) -- (-0.55+5,-2.75+2-8);
%\draw [thick](-0.55+5,-1.25-2) -- (-0.55+5,-2.75-2);

@@

@@

\section{Mean time}

\section{Discussion}

@@

\begin{figure}[!t]
 \vspace*{1\baselineskip}
\includegraphics[width=1\textwidth]{FractionPlot.pdf}
\caption{Plot of the fraction of the cases the population is rescue by standing genetic variation as a function of the fraction of initial cells which are aneuploid. The red vertical line highlights the the value of $f$ for which half the times the population is rescues by aneuploid cells while the pink line is our approximation \eqref{halfeqstandvar}. For this plot we have chosen the following parameters: $N=10^2, \lambda_w=1-10^{-2}, \lambda_a=1+10^{-2},\lambda_m=1+10^{-1},\mu_w=1,\mu_a=1,\mu_m=1.$}
\label{FractionPlot}
\end{figure}

>>>>

\draw [thick](0.15+10,-1.25+2-4) -- (0.15+10,-2.75+2-4);
\draw [thick](-0.25+10,-1.25+2-4) -- (-0.25+10,-2.75+2-4);
\draw [color=blue,very thick](-0.7+10,-1.25+2-4) -- (-0.7+10,-2.75+2-4);

\draw [thick](0.05+5,-1.25+2-8) -- (0.05+5,-2.75+2-8);
\draw [color=blue,very thick](-0.55+5,-1.25+2-8) -- (-0.55+5,-2.75+2-8);
%\draw [thick](-0.55+5,-1.25-2) -- (-0.55+5,-2.75-2);

@@

\begin{align}
\Delta_a^*=2vp_m+v+2\sqrt{vp_m\left(vp_m+\mu_a+v\right)}
\end{align}

@@

\section{Mean time}
%%%
\begin{figure}[!t]
 \vspace*{1\baselineskip}
\includegraphics[width=1\textwidth]{MeanTimeGrowthAneuploidyPlot.pdf}
\caption{}
\label{MeanTimeGrowthAneuploidyPlot}
\end{figure}
%%%

\begin{align}
\tau=\tau_1+\tau_2
\end{align}
inhomogeneus Poisson process
\begin{align}
\tau_1\sim \text{Poisson}\left(\frac{1}{up_aW\left(t\right)}\right)
\end{align}
where 
\begin{equation}
W\left(t\right)=W_0\e^{\Delta_wt}
\end{equation}
\begin{align}
f\left(\tau_1\right)=R\left(\tau_1\right)\e^{-\int_0^{\tau_1} R(t)\d t}
\end{align}
\begin{align*}
P\left(\tau_1<t\right)&=P\left(\tau_1<t|M\left(t\rightarrow\infty\right)\neq0\right)P\left(M\left(t\rightarrow\infty\right)\neq0\right)\\
&+P\left(\tau_1<t|M\left(t\rightarrow\infty\right)=0\right)P\left(M\left(t\rightarrow\infty\right)=0\right)\\
P\left(\tau_1<t\right)&=P\left(\tau_1<t|M\left(t\rightarrow\infty\right)\neq0\right)P\left(M\left(t\rightarrow\infty\right)\neq0\right)
\end{align*}
\begin{align}
P\left(\tau_1<t|M\left(t\rightarrow\infty\right)\neq0\right)=\frac{P\left(\tau_1<t\right)}{1-\left(1-p_w\right)^N}
\end{align}
where we have used
\begin{equation}
P\left(M\left(t\rightarrow\infty\right)\neq0\right)=1-\left(1-p_w\right)^N
\end{equation}
\begin{align}
f\left(\tau_1|M\left(t\rightarrow\infty\right)\neq0\right)=\frac{R\left(\tau_1\right)\e^{-\int_0^{\tau_1} R(t)\d t}}{1-\left(1-p_w\right)^N}
\end{align}
\begin{align}
\mathbb{E}\left[\tau_1\right]&=\frac{\int_0^\infty\e^{-\int_0^\tau R(t)\d t}\d\tau}{1-\left(1-p_w\right)^N}\\
&=\frac{\int_0^\infty\e^{-uNp_a\frac{\e^{\Delta_w\tau}-1}{\Delta_w}}\,\d\tau}{1-\left(1-p_w\right)^N}
\end{align}
\begin{align}
P_2\left(\tau_2\right)&=\int_0^\infty \d w P\left(W_l=w\right)\left[1-\e^{-vp_mW_l\left(\tau_2\right)}\right]\\
&=1-\mathbb{E}\left[\e^{-vp_mW_l\left(\tau_2\right)}\right]
\end{align}
\begin{align}
a_\pm=\frac{2-\delta-y\pm\sqrt{\left(\delta+y\right)^2-4y}}{2\left(1-\delta\right)}
\end{align}
\begin{align}
P_2\left(\tau_2\right)=\frac{\left(a_+-1\right)\left(1-a_-\right)\left(1-\e^{\left[-\left(1-\delta\right)\left(a_+-a_-\right)t\right]}\right)}{a_+-1+\left(1-a_-\right)\e^{\left[-\left(1-\delta\right)\left(a_+-a_-\right)t\right]}}
\end{align}
with $y=vp_m$ in $a_\pm$.
\begin{align}
\mathbb{E}\left[\tau_2\right]=\int_0^\infty\left(1-\frac{P_2\left(t\right)}{p_m}\right)\d t
\end{align}
\section{Discussion}

@@

\begin{figure}[!t]
 \vspace*{1\baselineskip}
\includegraphics[width=1\textwidth]{FractionPlot.pdf}
\caption{Plot of the fraction of the cases the population is rescue by standing genetic variation as a function of the fraction of initial cells which are aneuploid. The red vertical line highlights the the value of $f$ for which half the times the population is rescues by aneuploid cells while the pink line is our approximation \eqref{halfeqstandvar}. For this plot we have chosen the following parameters: $N=10^3, \lambda_w=1-10^{-2}, \lambda_a=1-10^{-4},\lambda_m=1+10^{-1},\mu_w=1,\mu_a=1,\mu_m=1.$}
\label{FractionPlot}
\end{figure}
